The Emergent Simplicity of Galaxy Size

### 2017-09-15 paper/galsize_paper.tex
Finished (very) rough overhaul of galaxy sample section

%%%%%%%%  galaxy properties  %%%%%%%%
\newcommand{\rhalf}{R_{1/2}}
\newcommand{\rhalftot}{R^{\rm tot}_{1/2}}
\newcommand{\rhalfdisk}{R_{1/2}^{\rm disk}}
\newcommand{\rhalfbulge}{R_{1/2}^{\rm bulge}}
\newcommand{\adisk}{A_{\rm disk}}
\newcommand{\abulge}{A_{\rm bulge}}
\newcommand{\alphadisk}{\alpha_{\rm disk}}
\newcommand{\alphabulge}{\alpha_{\rm bulge}}
\newcommand{\sigmarhalf}{\sigma_{\rm R_{1/2}}}
\newcommand{\rvir}{R_{\rm vir}}
\newcommand{\bt}{{\rm B/T}}
\newcommand{\mstar}{M_{\ast}}
\newcommand{\mstartot}{M^{\rm tot}_{\ast}}
\newcommand{\ssfr}{{\rm sSFR}}
\newcommand{\sfr}{{\rm SFR}}

@@

Our galaxy sample comes from Data Release 10 of the Sloan Digital Sky Survey \citep[SDSS][]{ahn_etal14}. We study the same $\mstar-$complete galaxy sample used in \citet{behroozi_etal15}, to which we refer the reader for details. Values for $\mstar$ and $\sfr$ are taken from the MPA-JHU value-added catalog \citep{kauffmann_etal03,brinchmann_etal04}.

We additionally use the catalog of SDSS galaxy profile decompositions provided by \citet{meert_etal15}. In the version of this catalog that we use, two-dimensional $r-$band profiles were fit with a two-component de Vaucouleurs + exponential profile. We cross-match our DR10 sample against the \citet{meert_etal15} catalog values of $\rhalftot,$ the half-light radius of the full (composite) profile, and $\bt,$ the $r-$band flux ratio of the de Vaucouleurs component to the total $r-$band flux. This cross-matching gives us a volume-limited galaxy sample with values of $\mstar,$ $\ssfr,$ $\rhalftot,$ and $\bt.$ Together with the present work, we make publicly available this cross-matched catalog, as well as code to generate the catalog from the original data files and load it into a python interpreter.

We calculate two-point clustering $\wproj$ of our SDSS galaxy sample using line-of-sight projection of $\pi_{\rm max}=20\mpc$ using the {\tt correl} program in {\tt UniverseMachine}. Our results in \S~\ref{sec:results} will give special focus on the dependence of $\wproj$ upon $\rhalftot.$ We will quantify this dependence in terms of {\em clustering ratios} of ``large" vs. ``small" galaxies, defined according to whether composite galaxy size is above or below $\langle\rhalftot\vert\mstartot\rangle,$ computed as the median of a sliding stellar mass window with a width of $N_{\rm gal}=1000.$

In all sections that follow, we will declutter our notation by referring to $\rhalftot$ as $\rhalf$ and $\mstartot$ as $\mstar.$ In all cases where we wish to refer to the half-light radius or stellar mass of a particular component, that component will explicitly appear in superscript. All numerical values of $\rhalf$ will be quoted in $\kpc,$ and all values of $\mstar$ and $\mhalo$ in $\msun,$ assuming $H_0=67.8~\kms\equiv100h~\kms,$ the best-fit value from \citet{planck15}. To scale stellar masses to ``$h=1$ units" \citep{croton13}, our numerically quoted values for $\mstar$ should be multiplied by a factor of $h^2,$ while our halo masses and distances should be multiplied by a factor of $h.$

@@

For central galaxies residing in host halos, we use the present-day virial radius; for satellite galaxies we use the virial radius at the time of accretion, using physical units of $\kpc.$ We model the composite size of model galaxies as
\beq
\rhalftot = (\bt)\rhalfbulge + (1-\bt)\rhalfdisk.
\eeq

@@

We reiterate an important distinction drawn in the formulation of our forward modeling approach. In the model described above, $\rhalfbulge$ and $\rhalfdisk$ are forward modeling quantities that we aim to constrain through SDSS observations of {\em composite} galaxy size $\rhalftot.$ In particular, both model and real galaxies will be categorized as ``small" or ``large`` according to whether $\rhalftot$ is above or below the median size for a galaxy with total stellar mass $\mstartot.$

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/random_bt_centrals_vs_satellites.pdf}
\caption{
{\bf Morphologies of centrals and satellites of the same halo mass.} We show the PDF of the bulge-to-total stellar mass ratio $\bt$ of central and satellite galaxies of the same halo mass $\mhalo=\mpeak.$ In the model, the PDF of $\bt$ is determined by the stellar mass $\mstar$ and specific star-formation rate $\ssfr$ statistically determine, with no residual dependence on the cosmic web. Satellite galaxies in the model are ``bulgier" than centrals of the same halo mass because satellites are more quiescent than centrals of the same stellar mass. Thus our baseline model ansatz is that the morphology-density relation is purely derived from the color-density relation. Since composite size $\rhalf$ in the model is given by $\rhalf\equiv\rhalftot=(\bt)\rhalfbulge + (1-\bt)\rhalfdisk,$ this ansatz in turn implies that smaller galaxies cluster more strongly relative to larger galaxies of the same stellar mass, as seen in the measurements shown in Fig.~\ref{fig:clustering_ratio_upshot}.
}
\label{fig:bt_censat}
\end{figure}
%-----------------------------------------------------------------------------------------------------

>>>>

%%%%%%%%  galaxy properties  %%%%%%%%
\newcommand{\rhalf}{R_{1/2}}
\newcommand{\rhalfdisk}{R_{1/2}^{\rm disk}}
\newcommand{\rhalfbulge}{R_{1/2}^{\rm bulge}}
\newcommand{\adisk}{A_{\rm disk}}
\newcommand{\abulge}{A_{\rm bulge}}
\newcommand{\alphadisk}{\alpha_{\rm disk}}
\newcommand{\alphabulge}{\alpha_{\rm bulge}}
\newcommand{\sigmarhalf}{\sigma_{\rm R_{1/2}}}
\newcommand{\rvir}{R_{\rm vir}}
\newcommand{\bt}{{\rm B/T}}
\newcommand{\mstar}{M_{\ast}}
\newcommand{\ssfr}{{\rm sSFR}}
\newcommand{\sfr}{{\rm SFR}}

@@

Our galaxy sample comes from the catalog of SDSS galaxy profile decompositions provided by \citet{meert_etal15}. This catalog is based on Data Release 10 of the Sloan Digital Sky Survey \citep[SDSS,][]{ahn_etal14}, with improvements to the photometry pipeline and light profile fitting methods \citep{vikram_etal10,bernardi_etal13,bernardi_etal14,meert_etal13}. In the version of this catalog that we use, two-dimensional $r-$band profiles were fit with a two-component de Vaucouleurs + exponential profile to determine the half-light radius $\rhalf.$ We apply the \citet{bell_etal03} mass-to-light ratio to the $r-$band flux and $g-r$ colors in this catalog to obtain an estimate for the total stellar mass $\mstar$ of every galaxy.

We calculate two-point clustering $\wproj$ of our SDSS galaxy sample using line-of-sight projection of $\pi_{\rm max}=20\mpc$ using the {\tt correl} program in {\tt UniverseMachine}. Our results in \S~\ref{sec:results} will give special focus on the dependence of $\wproj$ upon $\rhalf.$ We will quantify this dependence in terms of {\em clustering ratios} of ``large" vs. ``small" galaxies, defined according to whether composite galaxy size is above or below $\langle\rhalf\vert\mstar\rangle,$ computed as the median of a sliding stellar mass window with a width of $N_{\rm gal}=1000.$

All numerical values of $\rhalf$ will be quoted in $\kpc,$ and all values of $\mstar$ and $\mhalo$ in $\msun,$ assuming $H_0=67.8~\kms\equiv100h~\kms,$ the best-fit value from \citet{planck15}. To scale stellar masses to ``$h=1$ units" \citep{croton13}, our numerically quoted values for $\mstar$ should be multiplied by a factor of $h^2,$ while our halo masses and distances should be multiplied by a factor of $h.$

@@

For central galaxies residing in host halos, we use the present-day virial radius; for satellite galaxies we use the virial radius at the time of accretion, using physical units of $\kpc.$ We model the composite size of model galaxies as
\beq
\rhalf = (\bt)\rhalfbulge + (1-\bt)\rhalfdisk.
\eeq

@@

We reiterate an important distinction drawn in the formulation of our forward modeling approach. In the model described above, $\rhalfbulge$ and $\rhalfdisk$ are forward modeling quantities that we aim to constrain through SDSS observations of {\em composite} galaxy size $\rhalf.$ In particular, both model and real galaxies will be categorized as ``small" or ``large`` according to whether $\rhalf$ is above or below the median size for a galaxy with total stellar mass $\mstar.$

@@

%---------------------------------------------------------------------------------------------------
\begin{figure}
\centering
\includegraphics[width=8cm]{FIGS/random_bt_centrals_vs_satellites.pdf}
\caption{
{\bf Morphologies of centrals and satellites of the same halo mass.} We show the PDF of the bulge-to-total stellar mass ratio $\bt$ of central and satellite galaxies of the same halo mass $\mhalo=\mpeak.$ In the model, the PDF of $\bt$ is determined by the stellar mass $\mstar$ and specific star-formation rate $\ssfr$ statistically determine, with no residual dependence on the cosmic web. Satellite galaxies in the model are ``bulgier" than centrals of the same halo mass because satellites are more quiescent than centrals of the same stellar mass. Thus our baseline model ansatz is that the morphology-density relation is purely derived from the color-density relation. Since composite size $\rhalf$ in the model is given by $\rhalf\equiv\rhalf=(\bt)\rhalfbulge + (1-\bt)\rhalfdisk,$ this ansatz in turn implies that smaller galaxies cluster more strongly relative to larger galaxies of the same stellar mass, as seen in the measurements shown in Fig.~\ref{fig:clustering_ratio_upshot}.
}
\label{fig:bt_censat}
\end{figure}
%-----------------------------------------------------------------------------------------------------
